Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-06-26 WSSSPE2_report/WSSSPE2_review.tex
Section 4: shorten urls. Section 5: Comment a note.

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the
United Kingdom (UK). The SSI was founded to support the UK's research software
community by cultivating better, more sustainable research software to enable
world-class research. Chue Hong's keynote started by making the point
that the use of -- and reliance on -- software is pervasive in all areas of
world-leading research, showing examples from disciplines as diverse as
humanities and high-energy physics, quoting Kerstin Kleese Van Dam of the
Pacific Northwestern National Laboratory via a petition
campaign\footnote{\url{https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}}
at change.org, ``\emph{Today there are very few science areas left which do not
rely on IT and thus software for the majority of their research work. More
importantly key scientific advances in experimental and observational science
would have been impossible without better software.}'' He also cited Daniel
Katz, Software Infrastructure for Sustained Innovation (SI2) Program Director of
the NSF, ``\emph{Scientific discovery and innovation are advancing along
fundamentally new pathways opened by development of increasingly sophisticated
software. Software is an integral enabler of computation, experiment and theory,
and directly responsible for increased scientific productivity and enhancement
of researchers' capabilities.}''

Chue Hong drew attention to the issue that in the cyber-infrastructure and
high-performance community, hundreds of thousands
of researchers developing software are all too often disregarded or considered the long tail.
Actually, the numbers point to the fact that they are the mainstream. %and we
%(referring to people including the audience of the talk) are the elite minority.
He emphasized that software is no longer special; it is both essential to and
common in scientific research. A 2014
survey\footnote{\url{http://www.software.ac.uk/blog/2014-12-04-its-impossible-conduct-research-without-software-say-7-out-10-uk-researchers}}
conducted by the SSI polled researchers from 15 research-intensive UK
universities (406 respondents covering a representative range of funders,
disciplines, and seniority). The survey reported that 92\% respondents confirmed the use of research
software and 89\% affirmed that it would be impossible or difficult to conduct
research without software. Nevertheless the British research community is just
starting to understand the magnitude of the issue. Whilst many researchers make
use of software such as MATLAB, SPSS, and Excel, data from the aforementioned
SSI survey shows that over half (56\%) of respondents developed their own
research software (which equates to over 140,000 researchers if extrapolated
across the UK) and yet 71\% of all UK researchers had no formal software-development training, having
to rely on their own coding skills.

@@

\todo{writing led by Dan - other contributions welcome}

>>>>

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the
United Kingdom (UK). The SSI was founded to support the UK's research software
community by cultivating better, more sustainable software to enable
world-class research. Chue Hong's keynote started by making the point
that the use of -- and reliance on -- software is pervasive in all areas of
world-leading research, showing examples from disciplines as diverse as
humanities and high-energy physics, quoting Kerstin Kleese Van Dam of the
Pacific Northwestern National Laboratory via a petition
campaign\footnote{\url{http://tinyurl.com/nkn5tnv}}
%https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}}
at change.org, ``\emph{Today there are very few science areas left which do not
rely on IT and thus software for the majority of their research work. More
importantly key scientific advances in experimental and observational science
would have been impossible without better software.}'' He also cited Daniel
Katz, Software Infrastructure for Sustained Innovation (SI2) Program Director of
the NSF, ``\emph{Scientific discovery and innovation are advancing along
fundamentally new pathways opened by development of increasingly sophisticated
software. Software is an integral enabler of computation, experiment and theory,
and directly responsible for increased scientific productivity and enhancement
of researchers' capabilities.}''

Chue Hong drew attention to the issue that in the cyber-infrastructure and
high-performance community, hundreds of thousands
of researchers developing software are all too often disregarded or considered the long tail.
Actually, the numbers point to the fact that they are the mainstream. %and we
%(referring to people including the audience of the talk) are the elite minority.
He emphasized that software is no longer special; it is both essential to and
common in scientific research. A 2014
survey\footnote{\url{http://tinyurl.com/ooajs7m}}
%http://www.software.ac.uk/blog/2014-12-04-its-impossible-conduct-research-without-software-say-7-out-10-uk-researchers}}
conducted by the SSI polled researchers from 15 research-intensive UK
universities (406 respondents covering a representative range of funders,
disciplines, and seniority). The survey reported that 92\% respondents confirmed the use of research
software and 89\% affirmed that it would be impossible or difficult to conduct
research without software. Nevertheless the British research community is just
starting to understand the magnitude of the issue. Whilst many researchers make
use of software such as MATLAB, SPSS, and Excel, data from the aforementioned
SSI survey shows that over half (56\%) of respondents developed their own
research software (which equates to over 140,000 researchers if extrapolated
across the UK) and yet 71\% of all UK researchers had no formal software-development training, having
to rely on their own coding skills.

@@

%\todo{writing led by Dan - other contributions welcome}
